---
title: "Assess pairwise growth data"
author: "Brooke Anderson"
date: "7/13/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(ggplot2)
library(cowplot)
here::i_am("scripts/Assess_pairwise_growth_data.Rmd")
```

## load in data

```{r load data objects, echo = F}
d.18 <- readRDS(here("wrangled_data/all_pairwise.rds"))
```

## Pairwise data assessment

Are data normally distributed? -- run shapiro tests

```{r run shaprio tests for normality, echo = F, message = F, warning = F}
# define all conditions to run through
spec <- levels(d.18$species)
cond <- levels(d.18$condition)
pH <- levels(d.18$pH)

# are data normally distributed?
for (s in spec) {
  for (c in cond[which(cond != s)]) {
    for (p in pH) {
      for (d in c(3, 10, 21)) {
        dat <- d.18 %>% filter(species == s, condition == c, pH == p, day == d)
        if (sum(dat$total.CFUs) == 0 | nrow(dat) < 3) next
        if (shapiro.test(dat$total.CFUs)$"p.value" <= 0.05) {
          print(paste(s, "under condition", c, "at pH", p, "on day", d, "is not normal, with p value",
                     shapiro.test(dat$total.CFUs)$"p.value"))
        }
      }
    }
  }
}
```

Obviously, many samples are not normally distributed. Ah, biology. Anyway. Does log10 transformation help?

Are log-transformed data normally distributed? -- run shapiro tests

```{r run shaprio tests on log-transformed data, echo = F, message = F, warning = F}
# define all conditions to run through
spec <- levels(d.18$species)
cond <- levels(d.18$condition)
pH <- levels(d.18$pH)

# are data normally distributed?
for (s in spec) {
  for (c in cond[which(cond != s)]) {
    for (p in pH) {
      for (d in c(3, 10, 21)) {
        dat <- d.18 %>% filter(species == s, condition == c, pH == p, day == d) %>%
          mutate(log.CFUs = log10(total.CFUs))
        if (any(is.infinite(dat$log.CFUs)) | nrow(dat) < 3) next
        if (shapiro.test(dat$log.CFUs)$"p.value" <= 0.05) {
          print(paste(s, "CFUs under condition", c, "at pH", p, "on day", d, "is not normal, with p value",
                     shapiro.test(dat$log.CFUs)$"p.value"))
        }
      }
    }
  }
}
```


Better but still many samples are not normally distributed. 

Thinking ahead to the t-tests we'll want to do to compare coculture growth to alone growth, let's check how similar variances are between comparable samples. -- Levene's Test (of equality of variances for non-normally distributed data)
Start with raw colony-forming units:

```{r check variances for coculture vs. alone data, echo = F, message = F, warning = F}
# define all conditions to run through
spec <- levels(d.18$species)
cond <- levels(d.18$condition)[which(levels(d.18$condition) != "alone")]
pH <- levels(d.18$pH)

# for each species in each condition at each pH at each time, test for equal variance with alone sample
for (s in spec) {
  for (c in cond[which(cond != s)]) {
    for (p in pH) {
      for (d in c(3,10,21)) {
        # get cfu data for growth with partners
        dat.part <- d.18 %>% 
          filter(species == s, condition == c, pH == p, day == d)
        # get cfu data for growth alone
        dat.alone <- d.18 %>% 
          filter(species == s, condition == "alone", pH == p, day == d)
        # check similarity of variance using Levene's test
        levene.test <- car::leveneTest(total.CFUs ~ condition, data = rbind(dat.alone, dat.part))
        if (levene.test$"Pr(>F)"[1] <= 0.05) {
          print(paste(s, "CFUs alone versus with", c, "at pH", p, "on day", d, 
                      "do not have equal variance, with p value", levene.test$"Pr(>F)"[1]))
        }
      }
    }
  }
}
```

That's also not great. What happens to variances when CFU data is log-transformed? -- Levene's Test (of equality of variances for non-normally distributed data)

```{r check variances for log-transformed coculture vs. alone data, echo = F, message = F, warning = F}
# define all conditions to run through
spec <- levels(d.18$species)
cond <- levels(d.18$condition)[which(levels(d.18$condition) != "alone")]
pH <- levels(d.18$pH)

# for each species in each condition at each pH at each time, test for equal variance with alone sample
for (s in spec) {
  for (c in cond[which(cond != s)]) {
    for (p in pH) {
      for (d in c(3,10,21)) {
        # get cfu data for growth with partners
        dat.part <- d.18 %>% 
          filter(species == s, condition == c, pH == p, day == d) %>%
          mutate(log.CFUs = log10(total.CFUs))
        # get cfu data for growth alone
        dat.alone <- d.18 %>% 
          filter(species == s, condition == "alone", pH == p, day == d) %>%
          mutate(log.CFUs = log10(total.CFUs))
        # some samples will have 0's which become "-Inf" in log-transform, or all all 0's and can't be assessed
        if(!any(is.infinite(dat.part$log.CFUs))) {
          # check similarity of variance using Levene's test
          levene.test <- car::leveneTest(log.CFUs ~ condition, data = rbind(dat.alone, dat.part))
          if (levene.test$"Pr(>F)"[1] <= 0.05) {
            print(paste(s, "CFUs alone versus with", c, "at pH", p, "on day", d, 
                        "do not have equal variance, with p value", levene.test$"Pr(>F)"[1]))
          }
        }
      }
    }
  }
}
```

Just like for normality, that's a bit better. 

But the problem with log-transformed data is it turns real 0's into infinite numbers. And those can't be statistically assessed. So I lose lots of very real information when these zero-values are practically eliminated.

**So no log-transformation**, despite its help in normalizing sample distributions.
**Also use non-parametric tests.** Wilcoxon Rank-Sums tests compare the shift in two samples.

