---
title: "Plot Model Predictions"
author: "Brooke Anderson"
date: "7/15/2021"
output:
  pdf_document:
    fig_caption: yes
    fig_crop: no
  html_document:
    df_print: paged
header-includes: \usepackage{float}
geometry: margin=2cm
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(scales)
library(here)
here::i_am("scripts/plot_model_predictions.Rmd")
```

```{r load in plotting details, echo = F}
load(here("figures/plotting_objects.RData"))
```


## About the data

Collin took alone and pairwise CFU data for each species at day 21, assuming this was the "equilibrium" point (which would be where population size no longer changes, and dn/dt == 0), and used it to fit the Lotka-Volterra competition model to predict cell counts of each species in the complete community.

```{r load in data}
estimates <- rbind(read_csv(here("4_res/coefficient-estimates-pH5.csv")),
                   read_csv(here("4_res/coefficient-estimates-pH7.csv")) %>%
                     mutate(pH = 7)) %>%
  mutate(spec = recode(spec, "X135E" = "135E"), name = recode(name, "X135E" = "135E"))

predictions <- rbind(read_csv(here("4_res/predictions-from-pH5.csv")) %>%
                     pivot_longer(., cols = !c(rep, ph), names_to = "spec", values_to = "fit"),
                   read_csv(here("4_res/predictions-from-pH7.csv")) %>%
                     pivot_longer(., cols = !c(rep, ph), names_to = "spec", values_to = "fit")) %>%
  rename(pH = ph, replicate = rep) %>%
  mutate(spec = recode(spec, "X135E" = "135E")) %>%
  na.omit()

data <- read_csv(here("4_res/community-data-long.csv")) %>%
  mutate(spec = recode(spec, "X135E" = "135E"),
         fit = fit - 1)
colnames(data) <- c("replicate", "spec", "fit")

```

So I have: 

* actual community data,
* predictions of community CFUs from LV modeling of PW/alone growth data, and
* pairwise effect metrics from each pairwise partner species

Our main results are that:

* PW growth poorly predicts absolute community composition
* pH 7 PW data better predicts community composition than pH 5 data
* improved predictions from pH 7 pairwise data -- coming from Brevi/Brachy pairwise effect metrics?


Let's show these things:

PW growth poorly predicts absolute community composition

```{r by replicate actual vs. predicted cfus, echo = F, message = F, warning = F}

ggplot(data %>%
         mutate(spec = factor(spec, levels = names(names)[-c(8:9)])), 
       aes(x = spec, y = fit, group = replicate)) +
  geom_bar(aes(fill = spec), stat = "identity", color = "white",
           position = "dodge", width = 0.8, lwd = 1.3) +
  scale_fill_manual(values = species.palette, guide = "none") +
  scale_x_discrete(labels = names) +
  geom_point(data = predictions %>%
               mutate(fit.mod = ifelse(fit < 0, 0, fit),
                      pH = factor(pH, levels = c(5, 7)),
                      spec = factor(spec, levels = names(names)[-c(8:9)])) %>%
               filter(replicate %in% c(1,2,4,5)), 
             aes(x = spec, y = fit.mod, color = pH, group = replicate),
             position = position_dodge(width = 0.8)) +
  scale_color_manual(values = c("gray20", "red")) +
  scale_y_log10(limits = c(1, 1e10), breaks = c(1e2, 1e4, 1e6, 1e8, 1e0, 1e10),
                minor_breaks=trans_breaks("log10", function(x) 10^x),
                labels = trans_format("log10", label_math(10^.x)),
                expand = c(0,0)) +
  theme_bw() +
  labs(y = "colony-forming units",
       x = "community member",
       color = paste("media pH of\npairwise cocultures\nfor prediction"),
       subtitle = "bars show actual CFUs, points show replicate model predictions") +
  theme(axis.text.x = element_text(angle = 45, face = "italic", hjust = 1),
        plot.subtitle = element_text(size = 6))

ggsave(here("figures/suppl_fig_9.png"), device="png", width = 7, height = 4, units = "in")

  
```


```{r boxplot actual vs. predicted cfus, echo = F, message = F, warning = F}

ggplot(data %>%
         mutate(spec = factor(spec, levels = names(names)[-c(8:9)])), 
       aes(x = spec, y = fit)) +
  geom_boxplot(aes(fill = spec), width = .4, lwd = .4,
               outlier.shape = 21, outlier.size = 1) +
  scale_fill_manual(values = species.palette, guide = "none") +
  scale_x_discrete(labels = names) +
  geom_jitter(data = predictions %>%
               mutate(fit.mod = ifelse(fit < 0, 0, fit),
                      pH = factor(pH, levels = c(5, 7)),
                      spec = factor(spec, levels = names(names)[-c(8:9)])), 
             aes(x = spec, y = fit.mod, color = pH),
             width = .2, size = .8) +
  scale_color_manual(values = c("gray40", "red")) +
  scale_y_log10(limits = c(1, 1e10), breaks = c(1e2, 1e4, 1e6, 1e8, 1e0, 1e10),
                minor_breaks=trans_breaks("log10", function(x) 10^x),
                labels = trans_format("log10", label_math(10^.x))) +
  theme_bw() +
  labs(y = "colony-forming units",
       x = "community member",
       color = paste("media pH of\npairwise cocultures\nfor prediction"),
       subtitle = "boxplots show actual CFUs, points show replicate model predictions") +
  theme(axis.text.x = element_text(angle = 45, face = "italic", hjust = 1),
        plot.subtitle = element_text(size = 6))

  
```

Let's focus in on the deviations of predicted population size from actual CFUs when grown in a community:


```{r}

predictions.w <- predictions %>% 
  pivot_wider(names_from = "pH", names_prefix = "prediction_pH", 
              values_from = "fit")

comparisons <- full_join(data, # %>% mutate(fit = fit+1), # no infinite fold-changes, 
                         predictions.w, by = c("spec", "replicate")) %>%
  mutate(diff_ph5_pred = prediction_pH5 - fit,
         diff_ph7_pred = prediction_pH7 - fit,
         perc_ph5_pred = diff_ph5_pred / fit * 100,
         perc_ph7_pred = diff_ph7_pred / fit * 100,
         log_dev_ph5 = ifelse(perc_ph5_pred < 0, 
                              -log10(abs(perc_ph5_pred)),
                              log10(perc_ph5_pred)),
         log_dev_ph7 = ifelse(perc_ph7_pred < 0, 
                              -log10(abs(perc_ph7_pred)),
                              log10(perc_ph7_pred)))

p.deviations <- ggplot(comparisons %>% 
         pivot_longer(cols = c(log_dev_ph5, log_dev_ph7), values_to = "log_deviation",
                      names_to = "pH", names_prefix = "log_dev_ph") %>%
         mutate(spec = factor(spec, levels = names(names)[-c(8:9)])), 
       aes(x = spec, y = log_deviation, color = pH)) +
  scale_x_discrete(labels = names) +
  geom_jitter(size = .7, alpha = 0.7, width = .2) +
  scale_color_manual(values = c("gray30", "red")) +
  geom_hline(yintercept = 0, color = "gray40") +
  theme_bw() +
  labs(y = "log(scaled deviation)",
       x = "community member",
       color = paste("media pH of\npairwise cocultures\nfor prediction"),
       subtitle = "replicate log(%difference from actual) in small points") +
  theme(axis.text.x = element_text(angle = 45, face = "italic", hjust = 1),
        plot.subtitle = element_text(size = 6))

p.deviations + 
  stat_summary(geom = "point", fun = median, size = 2.5) +
  labs(subtitle = "replicate values in small points, median deviation in large points")


```


For *Diutina* (in that pH 7 predicts it to die off), *S. equorum*, *Brevibacterium*, and *Penicillium*, pH 7 PW data better predicts community composition than pH 5 data:

Can we look back into estimates and peg interactions that might be contributing extra to strong prediction deviations from actual CFUs?

Estimates should be interpreted as a per-capita replacement measurement: for every 1

```{r review estimates, echo = F}

estimates %>% 
  pivot_wider(values_from = c(est, se), names_from = pH) %>% 
  arrange(desc(abs(est_5))) %>%
  group_split(spec)
 
```

Some of these estimates aren't making sense -- e.g. at pH 5 for i = JB5, j = JB7, alpha_ij > 2e5? But neither of these species affect the other in pairwise coculture??

Let's generate a table that we can import into cytoscape to compare to actual effects:
```{r generate table for cytoscape, echo = F, message = F, warning = F}

# circular order of species nodes in network 
order <- c(`JB5` = 1, `JB7` = 2, `BC10` = 3, `BC9` = 4, `JBC` = 5, `JB370` = 6, `135E` = 7)

estimates %>% select(!se) %>%
  pivot_wider(names_from = name, values_from = est) %>%
  pivot_longer(cols = c("BC10", "BC9", "JB5", "JB7", `135E`, "JBC", "JB370"),
               names_to = "name", values_to = "est") %>%
  rename(species = spec, condition = name) %>%
  left_join(select(., species, K, pH) %>% 
              rename(condition = species, K.part = K) %>% 
              unique(),
            by = c("condition", "pH")) %>%
  filter(!is.na(est)) %>%  # only significant interactions
  # get magnitude of inhibitory & stimulatory effects on population size
  mutate(fc.est.per.source = est / K * 10e8,
         abs.fc.est.per.source = abs(fc.est.per.source), # convert to positive numbers for edge weighting
         # note that zero values are kept as zeros since cytoscape doesn't handle infinite measurements usefully
         # these zero values will be used to dictate dashing of edges for infinite fold-changes (i.e. non-detected coculture CFUs)
         `log.est.per.source` = ifelse(abs.fc.est.per.source != 0, 
                               log(abs.fc.est.per.source), # transform magnitude log10 and times two for visualization
                               3), # this value will determine the width of the edge in cytoscape; infinite values will be dashed with width of 3
         fc.est = est / K * K.part, # fold-change at day 21, accounting for source density 
         abs.fc.est = abs(fc.est),
         `log.est` = ifelse(abs.fc.est != 0, 
                               log(abs.fc.est), 
                               3),
         intx.sign = ifelse(est >= 0, "neg", "pos"), # indicate direction of pop size change
         target = paste0(intx.sign, species, "ph", pH), # combine details for target node
         source = paste0(intx.sign, condition, "ph", pH), # combine details for source node
         label = names[as.character(condition)], # provide source label
         order = order[as.character(condition)]) %>%
  select(target, source, label, order, intx.sign,
         est, abs.fc.est.per.source, `log.est.per.source`, 
         est, abs.fc.est, `log.est`) %>%
  write_csv(., here("wrangled_data/model_estimates_for_cytoscape.csv"))

```


```{r plot predictions as community relative abundance, echo = F, message = F, warning = F}
# get relative abundance data
pred.rel <- predictions %>%  
  mutate(value = paste0("pH ", pH),
         fit.mod = ifelse(fit < 0, 0, fit)) %>%
  group_by(value, replicate) %>% 
  mutate(rel.abund = fit.mod/sum(fit.mod)) %>%
  select(spec, value, replicate, rel.abund)
rel.abund <- rbind(pred.rel,
                   data %>%
                     group_by(replicate) %>% 
                     mutate(fit.mod = ifelse(fit < 0, 0, fit),
                            rel.abund = fit.mod/sum(fit.mod),
                            value = "actual") %>%
  select(spec, value, replicate, rel.abund)
)

rel.abund$value <- factor(rel.abund$value,
                          levels = c("actual", "pH 5", "pH 7"),
                          labels = c('italic("in vitro")~"community data"',
                                     '"prediction from pH 5 data"',
                                     '"prediction from pH 7 data"'))

# plot
ggplot(rel.abund %>% filter(replicate != 3)) + 
  geom_bar(aes(x = factor(replicate), y = rel.abund, 
               fill=factor(spec, 
                           levels = c("JB7","JB5","BC10","BC9", "JB370","JBC","135E"))), 
           stat = "identity", position = "fill", scale = "freex") +
  barp.theme + theme_bw() + 
  scale_y_continuous(labels = scales::percent, expand = c(0.005,0.005),
                     minor_breaks = c(0, .25, .5, .75, 1)) +
  labs(x = "replicate") +
  theme(legend.text = element_text(face = "italic"),
        panel.grid.major.x = element_blank()) +
  facet_grid(. ~ value, scales = "free_x", 
             labeller = label_parsed)

#ggsave(here("figures/rel_abundance_predicted_communities.png"), device="png", width = 7, height = 2.7, units = "in")

```

```{r plot predictions as community relative abundance, echo = F, message = F, warning = F}
# compile community predictions
pred.files <- list.files(here("4_res/"), pattern = "predictions-*", full.names = T)

l.pred <- lapply(pred.files, function(x) {
  read_csv(x) %>% 
    pivot_longer(cols = !c("rep", "ph"), names_to = "taxa", values_to = "cfus") %>%
    mutate(predictor = str_sub(basename(x), 13, -5),
           taxa = recode(taxa, "X135E" = "135E"))
})

pred <- do.call(rbind, l.pred)

# convert to relative abundance data
pred.rel <- pred %>%  
  mutate(fit.mod = ifelse(cfus < 0, 0, cfus)) %>%
  group_by(predictor, rep) %>% 
  mutate(rel.abund = fit.mod/sum(fit.mod)) %>% 
  select(taxa, predictor, rep, rel.abund)

rel.abund <- rbind(pred.rel %>%
                     group_by(taxa, predictor) %>%
                     summarize(rel.abund = mean(rel.abund)),
                   data %>%
                     group_by(replicate) %>% 
                     mutate(fit.mod = ifelse(fit < 0, 0, fit),
                            rel.abund = fit.mod/sum(fit.mod),
                            predictor = "actual",
                            taxa = spec) %>%
                     group_by(taxa, predictor) %>%
                     summarize(rel.abund = mean(rel.abund))) %>%
  mutate(predictor = factor(predictor,
         levels = c("actual", "from-pH5", "from-pH7",
                    "onlyJBC-135E-pH5", "onlyJBC-pH7"),
         labels = c('italic("in vitro")~"community data"',
                    '"prediction from all pH 5 data"',
                    '"prediction from all pH 7 data"',
                    'atop("prediction from pH 5, only",
                    italic("Dituina")~"&"~italic("Penicillium")~"effectors")',
                    '"prediction from pH 7, only"~italic("Penicillium")~"as an effector"')))

# %>% filter(predictor != "\"prediction from pH 7, only\"~italic(\"Penicillium\")~\"as an effector\"",
#                            predictor != "\"prediction from all pH 7 data\"" )

# plot
ggplot(rel.abund) + 
  geom_bar(aes(x = predictor, y = rel.abund, 
               fill=factor(taxa, 
                           levels = c("JB7","JB5","BC10","BC9", "JB370","JBC","135E"))), 
           stat = "identity", position = "fill", scale = "freex") +
  barp.theme + theme_bw() + 
  scale_y_continuous(labels = scales::percent, expand = c(0.005,0.005),
                     minor_breaks = c(0, .25, .5, .75, 1)) +
  scale_x_discrete(labels = function(x) parse(text=x)) +
  labs(x = NULL) +
  theme(legend.text = element_text(face = "italic"),
        panel.grid.major.x = element_blank(),
        axis.text.x = element_text(angle = 35, hjust = 1),
        plot.margin = margin(0.2, 0.2, 0.2, 1, "cm"))

ggsave(here("figures/rel_abundance_predicted_communities.png"), device="png", width = 5, height = 3.3, units = "in")

```

```{r}
all.pred <- rbind(data %>% mutate(predictor = "actual") %>% group_by(predictor, spec) %>%
        summarize(cfus = mean(fit)),
      pred %>% group_by(predictor, taxa) %>% 
        summarize(cfus = mean(cfus)) %>%
        rename(spec = taxa)) %>%
  mutate(cfus.mod = ifelse(cfus < 0, 0, cfus))

ggplot(all.pred %>%
         mutate(spec = factor(spec, levels = names(names)[-c(8:9)])), 
       aes(x = spec, y = cfus.mod)) +
  geom_bar(data = . %>% filter(predictor == "actual"),
           aes(fill = spec), stat = "identity", color = "white",
           position = "dodge", width = 0.8, lwd = 1.3) +
  scale_fill_manual(values = species.palette, guide = "none") +
  scale_x_discrete(labels = names) +
  geom_point(data = . %>% filter(predictor != "actual"),
             aes(shape = predictor)) +
  scale_y_log10(limits = c(1, 1e10), breaks = c(1e2, 1e4, 1e6, 1e8, 1e0, 1e10),
                minor_breaks=trans_breaks("log10", function(x) 10^x),
                labels = trans_format("log10", label_math(10^.x)),
                expand = c(0,0)) +
  theme_bw() +
  labs(y = "colony-forming units",
       x = "community member",
       color = paste("media pH of\npairwise cocultures\nfor prediction"),
       subtitle = "bars show actual CFUs, points show replicate model predictions") +
  theme(axis.text.x = element_text(angle = 45, face = "italic", hjust = 1),
        plot.subtitle = element_text(size = 6))
```


